### 2011-10-20 doc/manuscript.tex
I am still editting the Results sectioin with Ilan.

\pagestyle{myheadings}
%% ** EDIT HERE **

@@

\section*{Things To Do for submitting this to PLoS Pathogens}
\begin{enumerate}
\item A paragraph for briefing the methods of ClonalOrigin and Mowgli before
section \textbf{Results}.
\item We can add at least 50 words in section \textbf{Abstract}.
\item Figures, tables, and supporting information need to be ready.
\item Publish source codes at public domain.
\end{enumerate}

\section*{A list of our major and minor findings or observations}
\begin{enumerate}
\item We showed that recombining gene transfer from SPY to SDE was strong
enough.
\item A few evidences for the net directionality from SPY to SDE:
time-constrained gene transfer between SPY to SDE also showed the directionality
from SPY to SDE.
\item We showed that virulence genes were associated with horizontal gene
transfer. 
\item We found that recombining gene tranfer processes are involved in
membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
\item We found that horizontal gene transfer processes are involved in
transposases.
\item We provided evidence that recombination process affected more sites along
the core genome of SPY and SDE than mutation process.
\item We found a recombination hotspot at 200 kb from the origin of replication
along the genome of SDE1.
\item We showed that species-tree-topology-changing recombinant edges were more
reliably recovered than species-tree-topology-not-changing edges.
\item We showed that recombination intensity might be useful in finding genes
under high recombining gene transfer.
\item We found that SPY and SDE are more closely related than SDD and SDE in
about 10\% of the core genome.
\item Recombination intensity and number of recombining gene transfer events
were positively correlated.
\item We found reduction of numbers of genes after divergence of SDE and SPY.
\item We confirmed the species tree that relates the five genomes.
\item We listed several virulence genes associated with recombining and
horizontal gene transfer.
\end{enumerate}

\section*{Suggestions about the manuscript}
\begin{enumerate}
\item Ilan thought that a species tree must be shown at the outset of the
section \textbf{Results}.
\item Can selections after gene transfer be tested?
\end{enumerate}

@@

\texttt{We show a net directionality of recombining gene transfers from SPY to SDE. 
Describe our arguments of the net directionality. 
Number of recombinant edges.
Confidence about the numbers of recombinant edges.
Not all of genes experience recombining gene transfers from SPY to SDE.
Gene transfer counts from the parsimony-based approach; the power of the
parsimony-based approach may be limited.
Is this number of recombinant edges good enough for the argument?
Reflection of the previous analyses, 
heat maps of real data analysis, 
and supporting simulations.}

@@

We followed the pipeline and methods recommended by Didelot et al.
\cite{Didelot2010} to pursue a model-based study of recombining gene transfer. A
preliminary step in the pipeline involved estimating the species tree with
branch lengths (via \texttt{ClonalFrame}), and three population parameters of
mutation rate, recombination rate, and average recombinant tract length (via a
first phase of \texttt{ClonalOrigin}).  Because \textit{S.\ equi} ssp.\textit{\
equi} strain 4047 (SEE) diverged significantly from the others among the six
genomes (Table S\ref{tab:genome}), we could not obtain sufficiently long
multiple sequence alignments with it, and had to exclude it from the model-based
analysis.  The estimated species tree using the alignment of the five genomes
(Figure S\ref{fig:mauve}) had the expected topology \texttt{The figure 1 must be
nicer than one that now} (see Figure \ref{fig:tree5}). \texttt{ClonalFrame} also
estimated a relative strength of recombination compared to mutations
($r/m=4.66$), indicating potentially significant roles of recombination in the
evolutionary process of shaping these genomes; four times more sites had been
affected by recombination-driven substitutions than by mutation-driven
substitutions.  (Table S\ref{tab:clonalframe}).  Estimates of population
mutation rate ($\theta$), recombination rate ($\rho$), and average tract length
($\delta$) were obtained for the alignment blocks analyzed (Figure
S\ref{fig:scatter3}), and the median weighted over block lengths for each of the
three parameters was obtained as a global estimate (Table S\ref{tab:three} and
see \textbf{Materials and Methods}).  The local estimates typically seemed to be
moderately fluctuating around the global average, with the exception of what
appeared to be a recombination hotspot at a region 200 kilobase pairs from the
origin of replication of SDE1.

The estimated species tree together with the global estimates of the three
population parameters were used in the second phase of \texttt{ClonalOrigin},
where recombinant trees were sampled from the posterior distribution. One way to
summarize the posterior sample of recombinant trees was to count the topologies
of the local genealogies induced by it. Of the 105 possible rooted topologies
relatig five taxa, the dominant topology (66.7\%) was that of the species tree
(see \textbf{Materials and Methods} and Table S\ref{tab:Gene-tree-topologies}).
The second most frequently sampled topology (9.3\%) was the one in which SPY and
SDE were sister clades to each other with SDD being as an outgroup. Other
topologies were sampled with frequency less than 4\%. This gave initial
indication to the extent of recombining gene transfer  between SPY and SDE.

To gain a more fine-scale view of the main types of recombination scenarios, we
looked at the recombinant edges present in the posterior sample of recombinant
trees. Following the approach of Didelot et al. \cite{Didelot2010}, we
classified these edges by types defined by distinct pairs of donor and recipient
branches of the species tree (Figure \ref{fig:tree5}). For each edge type, we
recorded the average number of edges of that type present in the sampled
recombinant tree, and computed the ratio of this observed frequency to the prior
expected one as implied by the species tree (considering branch lengths and
their relative positions) and the three population parameters estimated in the
preliminary step of the model-based analysis.  Base 2 logarithms of these ratios
are depicted in the heatmap of Figure \ref{fig:Heatmap-of-recombination} (see
also Tables S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).

Among the possible recombinant edge types, we distinguished between ones that
preserve the species tree topology, and ones that alter it. We focused on the
topology-altering types, both because our main objective was to study
cross-clade gene transfer, and because our simulation study indicated that the
observed counts of topology-preserving recombinant edges were unreliably
inferred (see \textbf{Simulations}). The heatmap allowed to find recombination
scenarios that deviated from the expectation dictated by the probabilistic
model.  We envisioned two main potential causes for such deviations: (1) reduced
or increased opportunities for species to interact and exchange genetic
material; (2) selection acting on newly transferred genes.

We observed a reduced count of recombinant edges sampled between SDD and the
three branches of the SPY clade ($\sim$25\% of the expected count). This was
consistent with the fact that these species had few opportunities to interact
with each other, SPY being a human pathogen and SDD being strictly veterinary.
On the contrary, the number of sampled recombinant edges between SDD and
branches in the SDE clade seemed to be consistent with the prior, with a slight
excess for edges from SDE1 and SDE2 to SDD.  The greatest deviation of observed
counts from expected ones was observed for recombinant edges between the SPY and
SDE clades. It was concordant with the fact that these two species shared an
ecological niche; SDE was an opportunistic human pathogen. Interestingly, there
appeared to be much higher enrichment (up to four times more than expected) for
recombinant edges originating in the SPY clade and ending up in the SDE clade.
In a follow-up simulation study we show that this deviation from the expected
count was indeed highly significant, and so was the observed asymmetry in
pattern of recombination (see \textbf{Simulations}).  This net directionality of
recombination pattern from SPY to SDE was especially intriguing in light of the
hypothesis that SDE acquired human pathogenicity via gene transfer from SPY.

@@

Figure \ref{fig:Gene-duplication-loss} shows the inferred numbers of gene gain,
loss, duplication, and two different kinds of gene transfer events along
lineages of the tree that relates the six genomes. SPY had less genes than the
other species. All of the genomes were used to construct 2314 gene families.
The largest family consisted of 59 genes while nearly half of gene families
(1066) had exactly one gene per species, or six genes total (Figure
S\ref{fig:famsizes}).  We estimated a larger number of recombining gene transfer
events than that of horizontal gene transfer events; 944 and 294, respectively.
Duplications in extant lineages were more frequent than those in the other
lineages.  SDE lineages from the root appeared to have increased in the number
of genes while the common ancestor of SPY1 and SPY2 decreased in the number of
genes.  Whereas gene loss events were either comparable to or less than gene
gain ones along most of the lineages, it was not the case along the two internal
branches of SPY and SDE.  This indicated that SPY and SDE lineages had reduced
number of genes to specialize in their specific niches after each divergence
from their corresponding direct ancestors.

As in the model-based analysis, we attempted to see which type of horizontal or
recombining transfer events was most common in our inference.  We classified
events into types defined by donor-recipient branches in the species tree. We
recorded the number of inferred events of each type (Figures
S\ref{fig:hgt-heatmap} and S\ref{fig:mowgli-recomb-heatmap}). Unlike in the
model-based analysis, we could not normalize these counts by their expectation
because we did not have any model-based expectation for them, which restricted
our ability to interprest these counts. We did notice, however, a large number
of inferred recombining gene transfer events from the ancestral SPY branch to
the ancestral SDE branch, consistent with our observations from the model-based
analysis. Yet, we did not observed a similar pattern with horizontal gene
transfer while we observed a large number of horizontal transfers from the
ancesral SDE branch to SDD.

@@

We set out to seek categories of genes that were significantly enriched for
recombining or horizontal transfer events. For horizontal transfer, we used a
score based on a number of events inferred by the parsimony-based method, and
for recombining gene transfer, we used both the parsimony-based inferred counts
and a recombination intensity score based on the model-based analysis (see
\textbf{Materials and Methods}). First, we checked association of transfers with
virulence genes, which were provided from a previous study of Suzuki et al.\
\citet{Suzuki2011}.  This set of virulence genes was found to be significantly
associated with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but
not with recombining gene transfers. We also found a marginally significant
association between virulence genes and gene duplications (P-value $<7.13 \times
10^{-3}$).

We conducted similar tests for other functional categories of genes. Several
functional categories that were found to be associated with recombination
intensity were related with either membrane-bound molecules or aminoacyl tRNA
synthetases (Table \ref{tab:functional}). Consistent results were obtained for
the parsimony-based counts of recombining gene transfers (Table
S\ref{tab:go-events-recombining}).  Functional categories associated with
horizontal gene transfer were clearly different from those associated with
recombining gene transfer (Table \ref{tab:go-events}). The top ranked functional
categories in that list were related with transposases. These differences add to
our previous findings about the different roles these two types of transfer play
in the evolution of these species.

@@

{\bf [IG: MAYBE THERE IS A BETTER WAY TO DEMONSTRATE THE CORRELATION BETWEEN METHODS]}.

@@

Often computational and statistical analyses on genomes could benefit other
researchers when those resources were provided via Genome Browser.  We wished to
provide research community especially in SPY and SDE bacteria with results from
our analysis of gene transfer.  The resource shown in Figure S\ref{fig:ucsc} for
gene transfer was available at \url{http://strep-genome.bscb.cornell.edu}.  For
each of the five individual genomes as a reference genome several annotation
tracks were displayed: alignment blocks, recombination rate per site for each
block, local gene tree topologies, recombination intensity, locations of
virulence genes, posterior probability of recombining gene transfer. The genome
track of probability of recombining gene transfer allowed us to easily browse
the bacterial genome to locate regions that we might be interested in.  Using
the genome browser tracks we located virulence genes with notable posterior
recombination probability (Table S\ref{tab:virrecomb}).  About 10\% of the genes
comprised of virulence facters, about half of which were found in the core
genome of the five genome alignment \citep{Suzuki2011}.  By sorting virulence
genes by the number of horizontal gene transfer events we listed several genes
with high ranks (Table S\ref{tab:virhgt}).
The whole gene of  putative dipeptidase (SpyM3\_0465) was transferred, which was
a case for the finding that virulence genes tended to be transferred as a whole
unit \citep{Chan2009}.  Another example of whole gene transfer was superoxide
dismutase (SpyM3\_1071).  We investigated the unit of recombining gene transfer
by finding 58 genes for which a recombinant edge of certain types were sampled
with probability at least 0.9 by focusing on specific types of recombinant edges
(Table S\ref{tab:genes-transfer}).  In some cases, parts of a gene were inferred
to have been transferred, for instance, transfer from SPY1 to SDE1 was inferred
for 29\% of the aspartyl-tRNA synthetase gene. Recombinant tracts often were
shown to span two genes.  In other cases, the inferred transfer units spaned a
set of as much as eight adjacent genes: loci from SpyM3\_1518 to SpyM3\_1525.

@@

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

@@

%\section*{Figures\clearpage{}}

%=============================================================================
% Figure S2: Species tree
\begin{figure}
\includegraphics[scale=0.7]{figures/cornellf-3-tree}
\caption{\label{fig:tree5}The species tree of the five genomes inferred using
\texttt{ClonalFrame}.  SPY is the branch of a clade of \textit{Streptococcus
pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
and SD is the branch for the three individual genomes of \textit{S.\
dysgalactiae}. The scale is proportional to the expected number of mutations
given the Watterson's estimate of mutation rate and the tree.}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure 1. Heat map of recombining gene transfer from ClonalOrigin
\begin{figure}
\includegraphics[scale=0.45]{figures/heatmap-recedge}
\caption{\label{fig:Heatmap-of-recombination}Heatmap of recombination events
using the five genomes.  Each cell of the heat map represents the logarithm base
2 of the number of recombination events inferred relative to its expectation
under the prior, for a donor and recipient pair of branches. The donor is given by
the y-axis; the recipient on the x-axis.  The green colored cells are meaningless
because the recipient species tree branch of the cells are older than the donor
species branch.}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
% Figure 3. Gene duplication, loss, and transfer
\begin{figure}
\includegraphics[width=6in]{figures/strep-events}
\caption{\label{fig:Gene-duplication-loss} Distribution of inferred gene
duplication, loss, and transfer events across the phylogeny.  Gene counts for
each extant and ancestral species are given at each node of the phylogeny.  For
each branch we denote the number of gains G*, duplications D*, losses L*,
non-recombining transfers T*, and recombinations R*.  The non-recombining
transfer and recombination counts are indicated on the recipient branch (donors
are not indicated).}
\end{figure}
\clearpage{}%
%=============================================================================

@@

\caption{\label{fig:h3}
Simulations of the model-based approach for assessing the power of recovering
recombinant edges. Pairs of species tree branches are shown at x-axis labels.
Three intervals are displayed for each pair. Intervals with solid line with open
circle in the middle are for the logarithm of the fraction to base 2 of the average
number of recombinant edges from the ten simulated recombinant trees 
to the numbers of recombinant edges estimated from Simulation 2. 
Interverals with dotted line are for the logarithm of the fraction to base 2 of
the number of recombinant edges estimated from the real data
to the numbers of recombinant edges estimated from Simulation 2.
Intervals with solid line with closed circle in the middle are for the logarithm
of the fraction to base 2 of the number of recombinant edges estimated from the real data
to the numbers of recombinant edges estimated from Simulation 3. 
The horizontal lines are at the zero.
The intervals range from 5\% quantile to 95\% of quantile.}
\end{figure}
\clearpage{}%
%=============================================================================

@@

$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
\end{flushleft}

\end{table}
%\clearpage{}
%=============================================================================

>>>>

\pagestyle{myheadings}
%% ** EDIT HERE **
\markright{Gene Transfer using Streptococcus Genomes}

@@

% \section*{Things To Do for submitting this to PLoS Pathogens}
% \begin{enumerate}
% \item Legends for the main figures as a separate section
% \item We can add at least 50 words in section \textbf{Abstract}.
% \item Figures, tables, and supporting information need to be ready.
% \item Merge Genome Browser tracks to the official one.
% \item Publish source codes at public domain.
% \end{enumerate}
%
% \section*{A list of our major and minor findings or observations}
% \begin{enumerate}
% \item We showed that recombining gene transfer from SPY to SDE was strong
% enough.
% \item A few evidences for the net directionality from SPY to SDE:
% time-constrained gene transfer between SPY to SDE also showed the directionality
% from SPY to SDE.
% \item We showed that virulence genes were associated with horizontal gene
% transfer. 
% \item We found that recombining gene tranfer processes are involved in
% membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
% \item We found that horizontal gene transfer processes are involved in
% transposases.
% \item We provided evidence that recombination process affected more sites along
% the core genome of SPY and SDE than mutation process.
% \item We found a recombination hotspot at 200 kb from the origin of replication
% along the genome of SDE1.
% \item We showed that species-tree-topology-changing recombinant edges were more
% reliably recovered than species-tree-topology-not-changing edges.
% \item We showed that recombination intensity might be useful in finding genes
% under high recombining gene transfer.
% \item We found that SPY and SDE are more closely related than SDD and SDE in
% about 10\% of the core genome.
% \item Recombination intensity and number of recombining gene transfer events
% were positively correlated.
% \item We found reduction of numbers of genes after divergence of SDE and SPY.
% \item We confirmed the species tree that relates the five genomes.
% \item We listed several virulence genes associated with recombining and
% horizontal gene transfer.
% \end{enumerate}
% 
% \section*{Suggestions about the manuscript}
% \begin{enumerate}
% \item Ilan thought that a species tree must be shown at the outset of the
% section \textbf{Results}.
% \item Can selections after gene transfer be tested?
% \end{enumerate}

@@

% \texttt{We show a net directionality of recombining gene transfers from SPY to SDE. 
% Describe our arguments of the net directionality. 
% Number of recombinant edges.
% Confidence about the numbers of recombinant edges.
% Not all of genes experience recombining gene transfers from SPY to SDE.
% Gene transfer counts from the parsimony-based approach; the power of the
% parsimony-based approach may be limited.
% Is this number of recombinant edges good enough for the argument?
% Reflection of the previous analyses, 
% heat maps of real data analysis, 
% and supporting simulations.}

@@

We followed the pipeline and methods recommended by Didelot et al.
\cite{Didelot2010} to pursue a model-based study of recombining gene transfer. A
preliminary step in the pipeline involved estimating the species tree with
branch lengths via \texttt{ClonalFrame} \cite{Didelot2007}, and three population
parameters of mutation rate, recombination rate, and average recombinant tract
length via a first phase of \texttt{ClonalOrigin} \cite{Didelot2010}.  Because
\textit{S.\ equi} ssp.\textit{\ equi} strain 4047 (SEE) diverged significantly
from the others among the six genomes (Table S\ref{tab:genome}), we could not
obtain sufficiently long multiple sequence alignments with it, and had to
exclude it from the model-based analysis.  The estimated species tree using the
alignment of the five genomes (Figure S\ref{fig:mauve}) had the expected
topology \texttt{SC: The figure 1 must be nicer than one that now} (see Figure
\ref{fig:tree5}). \texttt{ClonalFrame} also estimated a relative strength of
recombination compared to mutations as $r/m=4.66$ meaning that four times more
sites had been affected by recombination-driven substitutions than by
mutation-driven substitutions (Table S\ref{tab:clonalframe}).  This indicated
potentially significant roles of recombination in the evolutionary process of
shaping these genomes.  Estimates of population mutation rate ($\theta$),
recombination rate ($\rho$), and average tract length ($\delta$) were obtained
for the alignment blocks analyzed (Figure S\ref{fig:scatter3}), and a median of
the estimates weighted over block lengths was obtained as a global estimate for
each of the three parameters (Table S\ref{tab:three}).  The local estimates
typically seemed to be moderately fluctuating around the global estimate, with
the exception of what appeared to be a recombination hotspot at a region 200
kilobase pairs from the origin of replication of SDE1.

The estimated species tree together with the global estimates of the three
population parameters were used to sample recombinant trees from the posterior
distribution in the second phase of \texttt{ClonalOrigin} \cite{Didelot2010}.
We summarized the posterior sample of recombinant trees by counting topologies
of the local genealogies induced by those recombinant trees for each of the
genomic positions (Table S\ref{tab:Gene-tree-topologies}).  Of the 105 possible
rooted topologies relatig five taxa, the dominant topology (66.7\%) was that of
the species tree.  The second most frequently sampled topology (9.3\%) was the
one in which SPY and SDE were sister clades to each other with SDD being as an
outgroup. Other topologies were sampled with frequency less than 4\%. This gave
initial indication to the extent of recombining gene transfer  between SPY and
SDE.

To gain a fine-scale view of the main types of recombination scenarios, we
looked at recombinant edges present in the posterior sample of recombinant
trees. Following the approach of Didelot et al. \cite{Didelot2010}, we
classified these edges by types defined by distinct pairs of donor and recipient
branches of the species tree (Figure \ref{fig:tree5}). For each edge type, we
averaged the number of edges present in the sampled recombinant tree, and
computed the ratio of this observed average frequency to a prior expected
frequency. The prior expected frequencies was computed by considering branch
lengths and their relative positions of the species tree, and the three
population parameter estimates in the preliminary step.  Base 2 logarithms of
these ratios are depicted in the heatmap of Figure
\ref{fig:Heatmap-of-recombination} (see also Tables S\ref{tab:heatmap} and
S\ref{tab:obsheatmap}).

Among the possible recombinant edge types, we can distinguish between ones that
preserve the species tree topology, and ones that alter it. We focused on 
topology-altering edge types both because our main objective was to study
cross-clade gene transfer, and because our simulation study indicated that the
observed counts of topology-preserving recombinant edges were unreliably
inferred (see \textbf{Simulations}). The heatmap allowed to find recombination
scenarios that deviated from the expectation dictated by the probabilistic
model.  We envisioned two main potential causes for such deviations: (1) reduced
or increased opportunities for species to interact and exchange genetic
material; (2) selection acting on newly transferred genes.

We observed reduction in the number of recombinant edges sampled between SDD and
the three branches of the SPY clade ($\sim$25\% of the expected count). This was
consistent with the fact that these species had few opportunities to interact
with each other, SPY being a human pathogen and SDD being a strictly veterinary
one.  On the contrary, the number of sampled recombinant edges between SDD and
branches in the SDE clade seemed to be consistent with the prior.  The greatest
deviation of observed counts from expected ones was in recombinant edges between
SPY and SDE clades. It was concordant with the fact that these two species
shared an ecological niche; SDE was an opportunistic human pathogen.
Interestingly, there appeared to be much higher enrichment (up to four times
more than expected) for recombinant edges originating in the SPY clade and
ending up in the SDE clade.  In a follow-up simulation study we concluded that
this deviation from the expected count was indeed highly significant, and so was
the observed asymmetry in pattern of recombination (see \textbf{Simulations}).
This net directionality of recombination pattern from SPY to SDE was especially
intriguing in light of the hypothesis that SDE acquired human pathogenicity via
gene transfer from SPY.

@@

Figure \ref{fig:Gene-duplication-loss} shows inferred numbers of gene gain,
loss, duplication, and two different kinds of gene transfer events along
lineages of the tree that relates the six genomes. SPY had less genes than the
other species. All of the six genomes were used to construct 2314 gene families.
The largest family consisted of 59 genes while nearly half of gene families
(1066) had exactly one gene per species, or six genes total (Figure
S\ref{fig:famsizes}).  We estimated a larger number of recombining gene transfer
events than that of horizontal gene transfer events; 944 and 294, respectively.
Duplications in extant lineages were more frequent than those in the other
lineages.  SDE lineages from the root appeared to have increased in the number
of genes while the common ancestor of SPY1 and SPY2 decreased in the number of
genes.  Whereas gene loss events were either comparable to or less than gene
gain events along most of the lineages, it was not the case along the two internal
branches of SPY and SDE.  This indicated that SPY and SDE lineages had reduced
number of genes to specialize in their specific niches after each divergence
from their corresponding direct ancestors.

As in the model-based analysis, we attempted to see which type of horizontal or
recombining transfer events was most common in our inference.  We classified
events into types defined by donor-recipient branches in the species tree. We
recorded the number of inferred events of each type (Figures
S\ref{fig:hgt-heatmap} and S\ref{fig:mowgli-recomb-heatmap}). Unlike in the
model-based analysis, we could not normalize these counts by their expectation
because we did not have any model-based expectation for them, which restricted
our ability to interpret these counts. We did notice, however, a large number of
inferred recombining gene transfer events from the ancestral SPY branch to the
ancestral SDE branch, consistent with our observations from the model-based
analysis. Yet, we did not observe any pattern of a net directional horizontal
gene transfer while we observed a large number of horizontal transfers from the
ancesral SDE branch to SDD.

@@

We set out to seek categories of genes that were significantly enriched for
recombining or horizontal transfer events. For horizontal transfer, we used a
score based on a number of events inferred by the parsimony-based method, and
for recombining gene transfer, we used both the parsimony-based inferred counts
and a recombination intensity score based on the model-based analysis (see
\textbf{Materials and Methods}). First, we checked association of transfers with
virulence genes, which were provided by a previous study of Suzuki et al.\
\citet{Suzuki2011}.  The set of virulence genes was found to be significantly
associated with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but
not with recombining gene transfers. We also found a marginally significant
association between virulence genes and gene duplications (P-value $<7.13 \times
10^{-3}$).

We conducted similar tests for other functional categories of genes. Several
functional categories that were found to be associated with recombination
intensity were related with either membrane-bound molecules or aminoacyl tRNA
synthetases (Table \ref{tab:functional}). Consistent results were obtained for
the parsimony-based counts of recombining gene transfers (Table
S\ref{tab:go-events-recombining}).  Functional categories associated with
horizontal gene transfer were clearly different from those associated with
recombining gene transfer (Table \ref{tab:go-events}). The top ranked functional
categories in that list were related with transposases. These differences added
to our previous findings about the different roles the two types of gene
transfer had played in the evolution of these species.

@@

@@

Often computational and statistical analyses on genomes could benefit other
researchers when those resources were available in a form of genome browser.  We wished to
provide research community especially in SPY and SDE bacteria with results from
our analysis of gene transfer.  The resource shown in Figure S\ref{fig:ucsc} for
gene transfer was available at \url{http://strep-genome.bscb.cornell.edu}.  For
each of the five individual genomes as a reference genome several annotation
tracks were displayed: alignment blocks, recombination rate per site for each
block, local gene tree topologies, recombination intensity, locations of
virulence genes, posterior probability of recombining gene transfer. The genome
track of probability of recombining gene transfer allowed us to easily browse
the bacterial genome to locate regions that we were interested in.  Using
the genome browser tracks we located virulence genes with notable posterior
recombination probability (Table S\ref{tab:virrecomb}).  About 10\% of the genes
comprised of virulence facters, about half of which were found in the core
genome of the five genome alignment \citep{Suzuki2011}.  After sorting virulence
genes by the number of horizontal gene transfer events we listed several genes
with high ranks (Table S\ref{tab:virhgt}).
The whole gene of  putative dipeptidase (SpyM3\_0465) was transferred, which was
a exemplary case for the finding \citep{Chan2009} that virulence genes were transferred as a whole
unit.  Another example of whole gene transfer was superoxide
dismutase (SpyM3\_1071).  We investigated the unit of recombining gene transfer
by finding 58 genes for which a recombinant edge of certain types were sampled
with probability at least 0.9 by focusing on specific types of recombinant edges
(Table S\ref{tab:genes-transfer}).  In some cases, parts of a gene were inferred
to have been transferred, for instance, transfer from SPY1 to SDE1 was inferred
for 29\% of the aspartyl-tRNA synthetase gene. Recombinant tracts often were
shown to span two genes.  In other cases, the inferred transfer units spaned a
set of as much as eight adjacent genes: loci from SpyM3\_1518 to SpyM3\_1525.

\subsection*{Simulation study and additional support for biased directionality in recombining gene transfer between SPY and SDE}

We conducted a series of experiments on simulated data to check the reliability of estimates obtained by ClonalOrigin. We found that the population parameters obtained in the first phase of ClonalOrigin are rather reliable, with the exception of the average recombinant tract length, which appeared to be over-estimated (this was observed also in \citep{Didelot2010}). Our experiments also show that the recombination intensity used in our association tests and traced in the genome browser tracks are reliably recovered by ClonalOrigin (Figure S\ref{fig:ri1}). See Materials and Methods for a more detailed report of these results. One issue we

 with a special focus on its ability to detect asymmetric patterns of recombination such as the one we found between SPY and SDE.

Numbers of 
recombinant edges from the simulated recombinant trees of Simulation 2 (see
\textbf{Supplementary Information} for the description of Simulation 2) were
compared with the corresponding numbers of recombinant edges that were estimated
from the simulated data set from Simulation 2 (see the solid-line intervals with
open circles in Figure \ref{fig:h3}).  We noted that tree-topology-changing
recombinant edges were more reliably recovered compared with
tree-topology-not-changing ones (Figure \ref{fig:h3}). All of the
tree-topology-not-changing recombinant edge types were over-estimated (Figure
\ref{fig:h3}B). The heat map of Figure \ref{fig:Heatmap-of-recombination} did
not show how confident of the values we could be. Comparison of numbers of
recombinant edges estimated from the real data set and the corresponding numbers
estimated from Simulation 2 showed the interval estimates for the heat map
values.  Dotted-line intervals from the red and blue colored cells in the heat
map were above and below the reference of the horizontal line at the zero on the
Y-axis, respectively (Figure \ref{fig:h3}).

@@

\section*{Figure Legends}

\begin{flushleft}

Figure \ref{fig:tree5}.\ \bf{The species tree of the five genomes inferred using
\texttt{ClonalFrame}.}  SPY is the branch of a clade of \textit{Streptococcus
pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
and SD is the branch for the three individual genomes of \textit{S.\
dysgalactiae}. The scale is proportional to the expected number of mutations
given the Watterson's estimate of mutation rate and the tree.

Figure \ref{fig:Heatmap-of-recombination}.\ \bf{Heatmap of recombination events 
using the five genomes.}
Each cell of the heat map represents the logarithm base
2 of the number of recombination events inferred relative to its expectation
under the prior, for a donor and recipient pair of branches. The donor is given by
the y-axis; the recipient on the x-axis.  The black colored cells are meaningless
because the recipient species tree branch of the cells are older than the donor
species branch.

Figure \ref{fig:Gene-duplication-loss}.\ \bf{Distribution of inferred gene 
duplication, loss, and transfer events across
the phylogeny.}  Gene counts for
each extant and ancestral species are given at each node of the phylogeny.  For
each branch we denote the number of gains G*, duplications D*, losses L*,
non-recombining transfers T*, and recombinations R*.  The non-recombining
transfer and recombination counts are indicated on the recipient branch (donors
are not indicated).

Figure \ref{fig:h3}.\ \bf{Simulations of the model-based approach for assessing 
the power of recovering recombinant edges.} 
Pairs of species tree branches are shown at x-axis labels.
Three intervals are displayed for each pair. Intervals with solid line with open
circle in the middle are for the logarithm of the fraction to base 2 of the average
number of recombinant edges from the ten simulated recombinant trees 
to the numbers of recombinant edges estimated from Simulation 2. 
Interverals with dotted line are for the logarithm of the fraction to base 2 of
the number of recombinant edges estimated from the real data
to the numbers of recombinant edges estimated from Simulation 2.
Intervals with solid line with closed circle in the middle are for the logarithm
of the fraction to base 2 of the number of recombinant edges estimated from the real data
to the numbers of recombinant edges estimated from Simulation 3. 
The horizontal lines are at the zero.
The intervals range from 5\% quantile to 95\% of quantile.

\end{flushleft}
\clearpage{}

%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

@@

%=============================================================================
% Figure 1: Species tree
\begin{figure}
\includegraphics[scale=0.7]{figures/cornellf-3-tree}
\caption{\label{fig:tree5}The species tree of the five genomes inferred using
\texttt{ClonalFrame}.  SPY is the branch of a clade of \textit{Streptococcus
pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
and SD is the branch for the three individual genomes of \textit{S.\
dysgalactiae}. The scale is proportional to the expected number of mutations
given the Watterson's estimate of mutation rate and the tree.}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure 2. Heat map of recombining gene transfer from ClonalOrigin
\begin{figure}
\includegraphics[scale=0.45]{figures/heatmap-recedge}
\caption{\label{fig:Heatmap-of-recombination}
\bf{Heatmap of recombination events using the five genomes.}
Each cell of the heat map represents the logarithm base
2 of the number of recombination events inferred relative to its expectation
under the prior, for a donor and recipient pair of branches. The donor is given by
the y-axis; the recipient on the x-axis.  The black colored cells are meaningless
because the recipient species tree branch of the cells are older than the donor
species branch.}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
% Figure 3. Gene duplication, loss, and transfer
\begin{figure}
\includegraphics[width=6in]{figures/strep-events}
\caption{\label{fig:Gene-duplication-loss} 
\bf{Distribution of inferred gene duplication, loss, and transfer events across
the phylogeny.}  Gene counts for
each extant and ancestral species are given at each node of the phylogeny.  For
each branch we denote the number of gains G*, duplications D*, losses L*,
non-recombining transfers T*, and recombinations R*.  The non-recombining
transfer and recombination counts are indicated on the recipient branch (donors
are not indicated).}
\end{figure}
\clearpage{}%
%=============================================================================

@@

\caption{\label{fig:h3}
\bf{Simulations of the model-based approach for assessing the power of recovering
recombinant edges.} Pairs of species tree branches are shown at x-axis labels.
Three intervals are displayed for each pair. Intervals with solid line with open
circle in the middle are for the logarithm of the fraction to base 2 of the average
number of recombinant edges from the ten simulated recombinant trees 
to the numbers of recombinant edges estimated from Simulation 2. 
Interverals with dotted line are for the logarithm of the fraction to base 2 of
the number of recombinant edges estimated from the real data
to the numbers of recombinant edges estimated from Simulation 2.
Intervals with solid line with closed circle in the middle are for the logarithm
of the fraction to base 2 of the number of recombinant edges estimated from the real data
to the numbers of recombinant edges estimated from Simulation 3. 
The horizontal lines are at the zero.
The intervals range from 5\% quantile to 95\% of quantile.}
\end{figure}
\clearpage{}%
%=============================================================================

@@

$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
\end{flushleft}
\label{tab:three}
\end{table}
%\clearpage{}
%=============================================================================
